Finiteness in Cubical Type Theory

### 2020-06-30 sections/finiteness-predicates.tex
done first proof in cardinal section

The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
  the same cardinality.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
  \end{agdalisting}
\end{lemma}
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
\end{agdalisting}
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-isSet}
\end{agdalisting}
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{alg}
\end{agdalisting}
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
\end{agdalisting}

\begin{proof}

  Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).
  Our task is to extract a natural number \(n : \mathbb{N}\) representing the
  cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
  equivalent to \(\AgdaDatatype{Fin}\;n\).

  Extracting the second component of the pair is trivial, as it itself is
  truncated.
  We will now focus on extracting the cardinality.

  Without the propositional truncation, \(\text{fst}\) would suffice for this
  task.
  Given that the pair is hidden under the truncation, then, we need a way to
  convert a function \(f : A \rightarrow B\) to \(g : \lVert A \rVert
  \rightarrow B\).
  This is precisely what eliminator \ref{elim-prop-coh} gives us.
  For our case, we need to show the following:
  \begin{equation}
    \frac{(n : \mathbb{N}) \; \; \; (p : \AgdaDatatype{Fin}\;n \simeq A) \; \; \;
          (m : \mathbb{N}) \; \; \; (q : \AgdaDatatype{Fin}(m) \simeq A)
        }{
          n \equiv m
        }
  \end{equation}
  Immediately we can construct the following term:
  \begin{equation}
    \begin{alignedat}{3}
      \AgdaDatatype{Fin}\;n & \simeq A && (p) \\
                      & \simeq \AgdaDatatype{Fin}(m) && (q)
    \end{alignedat}
  \end{equation}
  Given univalence we have \(\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)\),
  and the rest of our task is to prove:
  \begin{equation}
    \frac{\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)}{n \equiv m}
  \end{equation}

>>>>

The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
  the same cardinality.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
  \end{agdalisting}
\end{lemma}
\begin{proof}
  The high-level overview of our proof is as follows:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
  \end{agdalisting}
  It is the composition of two operations: first, with
  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
  proof of manifest bishop finiteness to a proof of equivalence with fin.

  Then we use the eliminator from Equation~\ref{elim-prop-coh} with three
  parameters.
  The first simply proves that that the output is a set:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-isSet}
  \end{agdalisting}
  The second is the function we apply to the truncated value:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{alg}
  \end{agdalisting}
  And the third is a proof that that function is itself coherently constant:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
  \end{agdalisting}

  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
  that \AgdaFunction{alg} returns the same value no matter its input.
  That output is a pair, the first component of which is the cardinality, and
  the second the truncated equivalence proof.
  The truncated proofs in the output are trivially equal by the truncation, so
  our obligation now has been reduced to:
  \begin{equation}
    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
    }{
      n\;\AgdaFunction{\ensuremath{\equiv}}\;m
    }
  \end{equation}
  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
  and the rest of our task is to prove:
  \begin{equation}
    \frac{\AgdaDatatype{Fin}\;n\;\AgdaDatatype{\ensuremath{\equiv}}\;\AgdaDatatype{Fin}\;m}{n \equiv m}
  \end{equation}
